Measuring precise radial velocities and cross-correlation function line-profile variations using a Skew Normal distribution

### 2018-05-21 manuscript.tex
Add files via upload

%\author{U. Simola \inst{1,2}
%	    \thanks{\email{umberto.simola@yale.edu}}
%	    \and X. Dumusque\inst{3}
%	    \thanks{Society in Science -- Branco Weiss Fellow (url: \url{http://www.society-in-science.org})}    
%	    \and Jessi Cisewski\inst{2}
%	    }
%
%\institute{Department of Statistical Sciences, University of Padua, Padua, Italy
%	      \and Department of Statistics and Data Science, Yale University, New Haven, CT, USA
%	      \and Observatoire de Gen\`eve, Universit\'e de Gen\`eve, 51 ch. des Maillettes, CH-1290 Versoix, Switzerland 
%	      }

@@

%\abstract
{\bf Abstract}\\
% Context, Aims, Methods, Results, Conclu (not mandatory)
Stellar activity is the main limitation to the detection of small-mass exoplanets using the radial-velocity (RV) technique. Stellar activity can be probed by measuring the cross-correlation function (CCF) shape variations as a function of time. Those variations are measured using the different moments of the CCF, therefore measuring with the best precision those moments is crucial to de-correlate exoplanet signals from spurious RV signals originating from stellar activity.

We propose here to measure those moments using a Skew Normal (SN) distribution, which compared to the Normal distribution generally used, naturally provides an extra parameter to model the CCF natural asymmetry induced by convective blueshift.

We analyze 5 stars with different activity levels and different signal-to-noise ratio levels. In each case, we compare the results obtained by fitting to the CCF respectively a Normal and a SN.  We also estimate rigorous errors for the different moments of the CCF using a bootstrap analysis.

The correlations between the RVs and the CCF asymmetry or RVs and the CCF width is always stronger when using the parameters derived from the SN in the case of real observations. Therefore the CCF asymmetry and width derived using a SN are more sensitive to stellar activity, which allows to probe with a better precision stellar rotational periods, in addition to characterize more precisely stellar activity signals. The precision on the RVs, derived using the median of the fitted SN distributions are on average $10\%$ smaller than the RV errors calculated on the mean of the Normal. In addition, the asymmetry parameter derived from the SN is on average $15\%$ better than the one calculated on the common Bisector Inverse Slope Span (BIS SPAN). We strongly encourage the use of the SN distribution rather than the Normal distribution, because this allows to retrieve in one single fit the different moments of the CCF, because the derived moments probe better stellar activity signals and because the standard errors on the RV and the the asymmetry are smaller.

@@

When working with radial velocities, the main limitation to the detection of small-mass exoplanets is not anymore the precision of the instruments used, but the different noises induce by the stars we are observing \citep[][]{Dumusque:2017aa}. Indeed, stellar oscillations, granulation phenomena and stellar activity all induce RV signals \citep[e.g.][]{Saar-1997b, Queloz-2001, Desort-2007, Dumusque-2011a, Dumusque-2016a} that are beyond the meter-per-second precision reached by the best high-resolution spectrographs. It is therefore mandatory to understand better stellar signals and to find ways to correct for them, if in the near future we want to detect or confirm an Earth-twin planet using the RV technique. This is even more true now that instrument like ESPRESSO \citep{Pepe-2014} and EXPRESS \citep{Fischer:2017aa} should have the stability to detect such signals. However, if solutions are not found to mitigate the impact of stellar activity, the detection or confirmation of potential Earth-twins will be extremely challenging.

@@

Among the different stellar signals we are aware of, the one that is the most difficult to characterize and to correct for is the signal induced by stellar activity. Stellar activity is responsible for creating magnetic regions on the surface of stars, and those regions change locally the temperature and the convection, inducing spurious RV variations \citep[e.g.][]{Meunier-2010a, Dumusque-2014b}. In theory, it should be easy to differentiate between the pure Doppler-shift induced by a planet, that will shift the entire stellar spectrum, and stellar activity that modifies the shape of spectral lines and by doing so create a spurious shift of the stellar spectrum \citep{Saar-1997b,Hatzes-2002,Kurster2003,Lindegren-2003,Desort-2007,Lagrange-2010,Meunier-2010a,Dumusque-2014b}. However, on quiet GKM dwarfs, the main target for precise RV measurements, stellar activity induce signals of a few \ms. This corresponds physically to variations smaller than 1/100th of a pixel on the detector.

To be able to measure such tiny variations, we average the information of all the line in the spectrum by cross correlating the stellar spectrum with a synthetic \citep[][]{Baranne-1996,Pepe-2002a} or an observed stellar template \citep[][]{Anglada-Escude-2012}.

The result of this operation gives us the cross-correlation function (CCF). To measure the Doppler-shift between different spectra, and therefore get the RV of a star as a function of time, we calculate the variations of the CCF barycenter. The barycenter is generally estimated by fitting a Normal distribution to the CCF, and taking its mean. Variation in line shape between different spectra, which indicate the presence of signals induced by stellar activity, is measured by analysing the different moments of the CCF. Generally, the width of the CCF is estimated using the FWHM of the fitted Normal distribution, and its asymmetry using the the bisector inverse slope span \citep[BIS SPAN,][]{Queloz-2001}.

@@

If a RV signal is induced by activity, generally a strong correlation will be observed between the RV and chromospheric activity indicators like \logrhk\,or H-$\alpha$ \citep[e.g.][]{Boisse-2009,Dumusque-2012,Robertson-2014}, but also between the RV and the FWHM of the CCF or its BIS SPAN \citep[][]{Queloz-2001,Boisse-2009,Queloz-2009,Dumusque-2016a}. It is therefore common now, that when fitting for a planetray signal, in addition to a Keplerian, the model includes in addition linear dependancies with the \logrhk, the FWHM and the BIS SPAN \citep[e.g.][]{Dumusque:2017aa,Feng:2017aa}. It is also common to add a Gaussian process to the model to account for the correlated noise induced by stellar activity. The hyperparameters of the Gaussian process are generally trained on the different activity indicators \citep[e.g.][]{Haywood-2014,Rajpaul-2015}. It is therefore essential for mitigating stellar activity to obtain activity indicators that are the most correlated to the RVs but also for which we can obtain the best precision.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Figueira indicators of stellar activity + other}
Several works have derived indicators that are more sensitive to line asymmetry than the BIS SPAN. In \citet{Boisse-2011}, the authors develop $V_{span}$ a new indicator to derive the CCF asymmetry, that is more sensitive than the BIS SPAN at low signal-to-noise ratio (SNR). \citet{Figueira-2013} studied the use of two new indicators, bi-Gauss and $V_{asy}$. The authors were able to show that when using bi-Gauss, the amplitude in asymmetry is 30\% larger, therefore allowing the detection of smaller-amplitude correlations with radial-velocity variations. They also demonstrated that $V_{asy}$ seems to be the best indicator of line asymmetry at high SNR, as its correlation with RV is more significant than any other asymmetry indictors.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Why using a SN distribution?}
In all the method described above, the RV and the FWHM is derived using a Normal distribution fitted to the CCF, and the asymmetry is estimated using another approach. In this paper we propose to use a Skew Normal (SN) distribution to derive at the same time the RV, the FWHM and the asymmetry of the CCF, as this function includes a skewness parameter \citep[][]{Azzalini1985}. In addition, we know that for solar-type stars and cooler dwarfs, the bisector of the CCF as a "C"-shape due to convective blueshift \citep[e.g.][]{Dravins-1981, Gray-2009}. Therefore, fitting the CCF using a model that naturally includes an asymmetry, like the SN distribution, should give more precise results.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Outline of paper}
The paper is organized as follow. In Sec.~\ref{sec:2} and in Sec.~\ref{sec:3} we introduce the SN distribution and describe its applicability for modeling the CCF,  showing that the SN distribution is a better representation of observed CCF than a Normal distribution, and study how the SN parameters relate to the RV, FWHM and BIS SPAN of the CCF. In Sec.~\ref{sec:3.1} we present a simple model to correct from stellar activity. In Sec.~\ref{sec:soap} we study the SN fit to the CCF using simulation coming from SOAP 2.0. In Sec.~\ref{sec:4}, we compare on real observations the sensitivity of the SN parameters to stellar activity with respect to other existing indicators. In Sec.~\ref{sec:5} we derive error bars for the different CCF parameters, and finally we discuss our results and conclude in Sec.~\ref{sec:discu} and Sec.~\ref{sec:conclu}.

@@

The Skew Normal (SN) distribution is a class of probability distributions which includes the Normal distribution as a special case \citep{Azzalini1985}. The SN distribution has, in addition to a location and a scale parameter analogous to the Normal distribution's mean and standard deviation, a third parameter which describes the asymmetry, or the skewness, of the distribution. Considering a random variable $Y\in \mathbb R$ (where $\mathbb R$ is the real line) which follows a SN distribution with location parameter $\xi \in \mathbb R$, scale parameter $\omega \in \mathbb R^{+}$ (i.e., the positive real line), and skewness parameter $\alpha \in \mathbb R$, its density at some value $Y = y$ can be written as 
\begin{equation} \label{def:snd_gen}
SN(y;\xi, \omega, \alpha) = \frac{2}{\omega} \phi\left(\frac{y-\xi}{\omega}\right) \Phi\left(\frac{\alpha(y-\xi)}{\omega}\right),
\end{equation}
where $\phi$ and $\Phi$ are respectively the density function and the distribution function of a \emph{standard} Normal distribution\footnote{A standard Normal distribution is a Normal distribution with a mean of 0 and a standard deviation of 1.}
 and $\alpha \in \mathbb R$ is the skewness parameter which quantifies the asymmetry of the SN.  We then write $Y \sim SN(\xi, \omega^{2}, \alpha)$ to mean that the random variable $Y$ follows the noted SN distribution.
Examples of SN densities under different skewness parameter values and the same location and scale parameters ($\xi = 0$ and $\omega = 1$) are displayed in Fig.~\ref{fig:SN.plot}.  A usual Normal distribution is the special case of the SN distribution when the skewness parameter, $\alpha$, is equal to 0.\footnote{This can be seen from Eq.~\eqref{def:snd_gen}.  If $\alpha = 0$ then $\Phi\left(\frac{\alpha(y-\xi)}{\omega}\right) = \Phi(0)$; this is the the probability a standard Normal random variable is less than or equal to 0, which is $0.5$.  The $0.5$ cancels with the $2$ in the denominator and what remains is the usual Normal density, $\frac{1}{\omega} \phi\left(\frac{y-\xi}{\omega}\right)$}
%
\begin{figure}[htbp]
   \centering
\includegraphics[height = 3in]{Skew_Normal_densities_jjck.pdf} 
   \caption{Density function of a random variable $Y \sim SN(\xi, \omega^{2}, \alpha)$ with location parameter $\xi = 0$, scale parameter $\omega = 1$ and different values of the skewness parameter $\alpha$ indicated by different colors and line types. Note that the solid black line has an $\alpha = 0$, making it a Normal distribution.}
   \label{fig:SN.plot}
\end{figure}
%
For reasons related to the interpretation of the parameters in Equation~\eqref{def:snd_gen} and computational issues with estimating $\alpha$ near 0, a different parametrization is used, which is referred to as the \emph{centered parametrization} (CP).  We will be using the CP in this work, which includes a mean parameter $\mu$, a variance parameter $\sigma^2$ and a skewness parameter $\gamma$.  In order to define the CP, we need to express the CP parameters $(\mu, \sigma^2, \gamma)$ as a function of the one used in the Equation~\eqref{def:snd_gen} with $(\xi, \omega^2, \alpha)$ by
%
\begin{equation} \label{eq:snd_cp}
\mu = \xi + \omega \beta, \quad \sigma^{2} = \omega^{2}(1-\beta^2), \quad \gamma = \frac{1}{2}(4-\pi) \beta^{3}\left(1-\beta^2\right)^{-3/2},
\end{equation}
%
where $\beta = \sqrt{\frac{2}{\pi}}\left(\frac{\alpha}{\sqrt{1+\alpha^2}}\right)$.

By using Eq.~\eqref{eq:snd_cp}, the new set of parameters $(\mu, \sigma^2, \gamma)$ provides a more clear interpretation of the behavior of the SN distribution. For the $\alpha$ values used in Fig.~\ref{fig:SN.plot}, the corresponding values of $\mu$, $\sigma^2$, $\gamma$ are displayed in Table~\ref{tab:cp_values}.  In particular, $\mu$ and $\sigma^2$ are the actual mean and variance of the distribution (rather than simply a location and scale parameter) and $\gamma$ becomes an index for evaluating the skewness of the SN.

@@

Since the CCF presents a natural asymmetry due the convective blueshift, the SN distribution should in principle better catch this aspect, together with other changes in asymmetry, respect the Normal fit. To initially check this assumption, we compared the CCF residuals after fitting a Normal and a SN distribution for 2 stars. The first star is Alpha Centauri b, whose CCFs have high signal-to-noise ratio (SNR). The second star is Corot-7, whose CCFs have low SNR. Fig.~\ref{fig:Residual.comparison} shows that the SN seems to be a slightly better model to explain the shape of the CCF, in particular as the SNR decreases.
%
\begin{figure*}[htbp]
   \centering
\includegraphics[height = 2.5in]{[1]HD128621Residuals.pdf} 
\includegraphics[height = 2.5in]{[1]LRa01_E2Residuals.pdf} 
   \caption{Comparison between the Normal (black circles) and the SN (red crosses) residuals using CCFs from the star Alpha Centauri b (left) and Corot-7 (right). When looking at the residuals corresponding to the tails of the CCF, the results derived by the two fits are comparable. However, when focusing on the center of the CCF, the SN fit leads to slightly better results for both the stars. Moreover, as the SNR decreases, the SN distribution shows smaller residuals respect the Normal ones.}
    \label{fig:Residual.comparison}
\end{figure*}
%

In the following of the paper, we define RV as the mean of the Normal distribution. Concerning the fit of the CCF using the SN, we present at first $2$ indicators that define the RV of the star: the mean of the SN, defined as RV mean SN and the median of the SN, defined as RV median SN (i.e. realling Eq.\eqref{eq:snmed}, $m$ = RV median SN). We will discuss advantages and limits for both these choices in Section \ref{sec:4} and Section \ref{sec:5}. For the width of the CCF, we use the FWHM of the Normal, which is $2\sqrt(2\ln2)\sigma$. The width of the SN, SN FWHM, is defined in the same way\footnote{Note that SN FWHM does not correspond to the width of the SN distribution at half maximum like in the Normal case.}. Being a Normal distribution symmetric, there is not such a parameter that evaluates the asymmetry of the distribution, so the BIS SPAN is used. The BIS SPAN will be compared to the asymmetric parameter $\gamma$ of the SN, named SN GAMMA. To test the strength of the correlation between the RVs and the different indicators for stellar activity, we calculated the Pearson correlation coefficient, $R$.
A $p-$value for the statistical test having null hypothesis $H_{0}: R=0$ (i.e., no correlation) is provided, along with a $95\,\%$ confidence interval for $R$.

@@

To test the strength of the correlation between the estimated RV's and the different indicators used to evaluate stellar activity, we calculated the Pearson correlation coefficient, which in its general form is defined as:
%
\begin{equation}
R (x,y)= \frac{cov(x,y)}{\sigma(x),\sigma(y)},
\label{eq:Pearson:corr}
\end{equation}
where $x$ and $y$ are two quantitative variables, $cov(x,y)$ indicates the covariance between $x$ and $y$, and $\sigma(x)$ and $\sigma(y)$ represent their standard deviations. 
A $p-$value for the statistical test having null hypothesis $H_{0}: R=0$ is provided, along with a $95\,\%$ confidence interval for $R$.
%
Because of the spurious variations in RV's caused by active regions, a crucial step of the entire analysis is to correct the initially estimated RV's in order to retrieve a set of RV's containing ideally only the possibly pure doppler shift coming from the exoplanets. In order to reach this goal, we propose to consider a linear combination of the RV's having as covariates the amplitude parameter A, the BIS SPAN, the FWHM and the interaction between the BIS SPAN and the FWHM (or $\gamma$ and SN FWHM and the interaction between $\gamma$ and SN FWHM in the SN case), as shown in Eq.~\eqref{eq:RV:correction}.
%
\begin{equation}
RV_{\text{stellar activity}}= \beta_{0} + \beta_{1} A + \beta_{2} \gamma + \beta_{3} \text{SN FWHM} + \beta_{4} (\gamma  \text{SN FWHM})+\epsilon,
\label{eq:RV:correction}
\end{equation}
%
where $\beta_{0}$ is the intercept and $\epsilon$ is the vector of the errors with mean equal to $0$ and covariance matrix equal to $\sigma^{2}I$ ($I$ defined as the identity matrix).

While the parameters $\gamma$ (or BIS SPAN) and SN FWHM (or FWHM) have already been introduced, in Eq.~\eqref{eq:RV:correction} we propose also two include the amplitude parameter $A$ and a variable that quantifies the interaction between $\gamma$ (or BIS SPAN) and SN FWHM (or FWHM). The inclusion of the amplitude parameter A is justified by the expected behavior of the CCF in presence of active regions. Fig.~\ref{fig:spotA} show that the presence for instance of a spot produces also changes on the amplitude of the CCF and not only on its asymmetry or width.
%
\begin{figure*}[htbp]
   \centering
\includegraphics[height = 4in]{FTS_solar_CCFs_whyWeAddAlsoPARAM_A.pdf} 
   \caption{The presence of a spot on the photosphere of a star produces changes on the amplitude of the CCF and not only on its asymmetry or width.}
    \label{fig:spotA}
\end{figure*}
%
The reasons for including in Eq.~\eqref{eq:RV:correction} a variable that quantifies the interaction between $\gamma$ (or BIS SPAN) and SN FWHM (or FWHM) will be clarified in the next Section and are related to a first attempt to disentangle spots from faculae.

>>>>

%\author{U. Simola \inst{1,2}
%	    \thanks{\email{umberto.simola@yale.edu}}
%	    \and X. Dumusque\inst{3}
%	    \thanks{Society in Science -- Branco Weiss Fellow (url: \url{http://www.society-in-science.org})}    
%	    \and Jessi Cisewski\inst{2}
%	    }
%
%\institute{Department of Mathematic and Statistics, University of Helsinki, Helsinki, Finland
%	      \and Department of Statistics and Data Science, Yale University, New Haven, CT, USA
%	      \and Observatoire de Gen\`eve, Universit\'e de Gen\`eve, 51 ch. des Maillettes, CH-1290 Versoix, Switzerland 
%	      }

@@

%\abstract
{\bf Abstract}\\
% Context, Aims, Methods, Results, Conclu (not mandatory)
Stellar activity is the main limitation to the detection of small-mass exoplanets using the radial-velocity (RV) technique. Stellar activity can be probed by measuring variations in the shape of the cross-correlation function (CCF) as a function of time. Those variations are measured using the different moments of the CCF, therefore measuring with the best precision those moments is crucial to de-correlate exoplanet signals from spurious RV signals originating from stellar activity.

We propose here to measure those moments using a Skew Normal (SN) distribution, which compared to the Normal distribution generally used, naturally provides an extra parameter to model the natural asymmetry of the CCF induced by convective blueshift.

We analyze 5 stars with different activity levels and different signal-to-noise ratio (SNR) levels. In each case, we compare the results obtained by fitting to the CCF respectively a Normal and a SN.  We also estimate rigorous errors for the different moments of the CCF using a bootstrap analysis.

The correlation between the RV's and the CCF asymmetry or the correlation between the RV's and the CCF width are always stronger when using the parameters derived from the SN in the case of real observations. Therefore the CCF asymmetry and the CCF width derived using a SN are more sensitive to stellar activity, which allows to probe with a better precision stellar rotational periods, in addition to characterize more precisely stellar activity signals. The precision on the estimated set of RV's, derived using the median of the fitted SN distributions are on average $10\%$ smaller than the uncertainties calculated on the mean of the Normal. In addition, the asymmetry parameter derived from the SN is on average $15\%$ better than the one calculated on the common Bisector Inverse Slope Span (BIS SPAN). We strongly encourage the use of the SN distribution rather than the Normal distribution, because this allows us to retrieve in one single fit the different moments of the CCF, because the derived moments better catch stellar activity signals and finally because the standard errors on the RV's and the the asymmetry parameter are smaller than the one estimated with the Normal fit.

@@

When working with radial velocities (RV's), the main limitation to the detection of small-mass exoplanets is not anymore the precision of the instruments used, but the different noises induced by the stars we are observing \citep[][]{Dumusque:2017aa}. Indeed, stellar oscillations, granulation phenomena and stellar activity all induce RV signals \citep[e.g.][]{Saar-1997b, Queloz-2001, Desort-2007, Dumusque-2011a, Dumusque-2016a} that are beyond the meter-per-second precision reached by the best high-resolution spectrographs. It is therefore mandatory to understand better stellar signals and to find ways to correct for them, if in the near future we want to detect or confirm an Earth-twin planet using the RV technique. This is even more true now that instrument like ESPRESSO \citep{Pepe-2014} and EXPRESS \citep{Fischer:2017aa} should have the stability to detect such signals. However, if solutions are not found to mitigate the impact of stellar activity, the detection or confirmation of potential Earth-twins will be extremely challenging.

@@

Among the different stellar signals we are aware of, the one that is the most difficult to characterize and to correct for is the signal induced by stellar activity. Stellar activity is responsible for creating magnetic regions on the surface of stars, and those regions change locally the temperature and the convection, inducing spurious RV's variations \citep[e.g.][]{Meunier-2010a, Dumusque-2014b}. In theory, it should be easy to differentiate between the pure Doppler-shift induced by a planet, that will shift the entire stellar spectrum, and stellar activity that modifies the shape of spectral lines and by doing so create a spurious shift of the stellar spectrum \citep{Saar-1997b,Hatzes-2002,Kurster2003,Lindegren-2003,Desort-2007,Lagrange-2010,Meunier-2010a,Dumusque-2014b}. However, on quiet GKM dwarfs, the main target for precise RV's measurements, stellar activity induce signals of a few \ms. This corresponds physically to variations smaller than 1/100th of a pixel on the detector.

To be able to measure such tiny variations, we average the information of all the line in the spectrum by cross correlating the stellar spectrum with a synthetic \citep[][]{Baranne-1996,Pepe-2002a} or an observed stellar template \citep[][]{Anglada-Escude-2012}. The result of this operation gives us the cross-correlation function (CCF). To measure the Doppler-shift between different spectra, and therefore get the RV's of a star as a function of time, we calculate the variations of the CCF barycenter. The barycenter is generally estimated by fitting a Normal distribution to the CCF, and taking its mean. Variations in line shape between different spectra, which indicate the presence of signals induced by stellar activity, is measured by analysing the different moments of the CCF. Generally, the width of the CCF is estimated using the FWHM of the fitted Normal distribution, and its asymmetry using the the bisector inverse slope span \citep[BIS SPAN,][]{Queloz-2001}.

@@

If a RV signal is induced by activity, generally a strong correlation will be observed between the RV and chromospheric activity indicators like \logrhk\,or H-$\alpha$ \citep[e.g.][]{Boisse-2009,Dumusque-2012,Robertson-2014}, but also between the RV and the FWHM of the CCF or its BIS SPAN \citep[][]{Queloz-2001,Boisse-2009,Queloz-2009,Dumusque-2016a}. It is therefore common now, that when fitting for a planetray signal, in addition to a Keplerian, the model includes in addition linear dependancies with the \logrhk, the FWHM and the BIS SPAN \citep[e.g.][]{Dumusque:2017aa,Feng:2017aa}. It is also common to add a Gaussian process to the model to account for the correlated noise induced by stellar activity. The hyperparameters of the Gaussian process are generally trained on the different activity indicators \citep[e.g.][]{Haywood-2014,Rajpaul-2015}. It is therefore essential for mitigating stellar activity to obtain activity indicators that are the most correlated to the RV's but also for which we can obtain the best precision.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Figueira indicators of stellar activity + other}
Several works have derived indicators that are more sensitive to line asymmetry than the BIS SPAN. In \citet{Boisse-2011}, the authors develop $V_{span}$, a new indicator to derive the CCF asymmetry that is more sensitive than the BIS SPAN at low signal-to-noise ratio (SNR). \citet{Figueira-2013} studied the use of two new indicators, bi-Gauss and $V_{asy}$. The authors were able to show that when using bi-Gauss, the amplitude in asymmetry is 30\% larger, therefore allowing the detection of smaller-amplitude correlations with RV's variations. They also demonstrated that $V_{asy}$ seems to be the best indicator of line asymmetry at high SNR, as its correlation with RV is more significant than any other asymmetry indictors.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Why using a SN distribution?}
In all the method described above, the RV and the FWHM are derived using a Normal distribution fitted to the CCF, and the asymmetry is estimated using another approach.

In this paper we propose to use a Skew Normal (SN) distribution to derive at the same time the RV, the FWHM and the asymmetry of the CCF, as this function includes a skewness parameter \citep[][]{Azzalini1985}. In addition, we know that for solar-type stars and cooler dwarfs, the bisector of the CCF as a "C"-shape due to convective blueshift \citep[e.g.][]{Dravins-1981, Gray-2009}. Therefore, fitting the CCF using a model that naturally includes an asymmetry, like the SN distribution, should give more precise results.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Outline of paper}
The paper is organized as follow. In Sec.~\ref{sec:2} we introduce the SN distribution and describe its applicability for modeling the CCF,  showing that the SN distribution is a better representation of observed CCF than a Normal distribution, and study how the SN parameters relate to the RV, FWHM and BIS SPAN of the CCF. In Sec.~\ref{sec:31} we present a simple model to correct from stellar activity which extends the ones proposed by \citet[e.g.][]{Dumusque:2017aa,Feng:2017aa}. In Sec.~\ref{sec:soap} we study the SN fit to the CCF using simulations coming from SOAP 2.0. In Sec.~\ref{sec:4}, we compare on real observations the sensitivity of the SN parameters to stellar activity with respect to other existing indicators. In Sec.~\ref{sec:5} we derive error bars for the different CCF moments and finally we discuss our results and conclude in Sec.~\ref{sec:discu} and Sec.~\ref{sec:conclu}.

@@

The Skew Normal (SN) distribution is a class of probability distributions which includes the Normal distribution as a special case \citep{Azzalini1985}. The SN distribution has, in addition to a location and a scale parameter analogous to the Normal distribution's mean and standard deviation, a third parameter which describes the asymmetry, or the skewness, of the distribution. Considering a random variable $Y\in \mathbb R$ (where $\mathbb R$ is the real line) which follows a SN distribution with location parameter $\xi \in \mathbb R$, scale parameter $\omega \in \mathbb R^{+}$ (i.e., the positive real line), and skewness parameter $\alpha \in \mathbb R$, its density at some value $Y = y$ can be written as 
\begin{equation} \label{def:snd_gen}
SN(y;\xi, \omega, \alpha) = \frac{2}{\omega} \phi\left(\frac{y-\xi}{\omega}\right) \Phi\left(\frac{\alpha(y-\xi)}{\omega}\right),
\end{equation}
where $\phi$ and $\Phi$ are respectively the density function and the distribution function of a \emph{standard} Normal distribution\footnote{A standard Normal distribution is a Normal distribution with a mean of 0 and a standard deviation of 1.}
 and $\alpha \in \mathbb R$ is the skewness parameter which quantifies the asymmetry of the SN.  We then write $Y \sim SN(\xi, \omega^{2}, \alpha)$ to mean that the random variable $Y$ follows the noted SN distribution.
Examples of SN densities under different skewness parameter values and the same location and scale parameters ($\xi = 0$ and $\omega = 1$) are displayed in Fig.~\ref{fig:SN.plot}.  A usual Normal distribution is the special case of the SN distribution when the skewness parameter, $\alpha$, is equal to 0.\footnote{This can be seen from Eq.~\eqref{def:snd_gen}.  If $\alpha = 0$ then $\Phi\left(\frac{\alpha(y-\xi)}{\omega}\right) = \Phi(0)$; this is the the probability a standard Normal random variable is less than or equal to 0, which is $0.5$.  The $0.5$ cancels with the $2$ in the denominator and what remains is the usual Normal density, $\frac{1}{\omega} \phi\left(\frac{y-\xi}{\omega}\right)$}
%
\begin{figure}[htbp]
   \centering
\includegraphics[height = 3in]{Skew_Normal_densities_jjck.pdf} 
   \caption{Density function of a random variable $Y \sim SN(\xi, \omega^{2}, \alpha)$ with location parameter $\xi = 0$, scale parameter $\omega = 1$ and different values of the skewness parameter $\alpha$ indicated by different colors and line types. Note that the solid black line has an $\alpha = 0$, making it a Normal distribution.}
   \label{fig:SN.plot}
\end{figure}
%
For reasons related to the interpretation of the parameters in Eq.~\eqref{def:snd_gen} and computational issues with estimating $\alpha$ near 0, a different parametrization is used, which is referred to as the \emph{centered parametrization} (CP).  We will be using the CP in this work, which includes a mean parameter $\mu$, a variance parameter $\sigma^2$ and a skewness parameter $\gamma$.  In order to define the CP, we need to express the CP parameters $(\mu, \sigma^2, \gamma)$ as a function of the one used in the Equation~\eqref{def:snd_gen} with $(\xi, \omega^2, \alpha)$ by
%
\begin{equation} \label{eq:snd_cp}
\mu = \xi + \omega \beta, \quad \sigma^{2} = \omega^{2}(1-\beta^2), \quad \gamma = \frac{1}{2}(4-\pi) \beta^{3}\left(1-\beta^2\right)^{-3/2},
\end{equation}
%
where $\beta = \sqrt{\frac{2}{\pi}}\left(\frac{\alpha}{\sqrt{1+\alpha^2}}\right)$.

By using Eq.~\eqref{eq:snd_cp}, the new set of parameters $(\mu, \sigma^2, \gamma)$ provides a more clear interpretation of the behavior of the SN distribution. For the $\alpha$ values used in Fig.~\ref{fig:SN.plot}, the corresponding values of ($\mu$, $\sigma^2$, $\gamma$) are displayed in Table~\ref{tab:cp_values}.  In particular, $\mu$ and $\sigma^2$ are the actual mean and variance of the distribution (rather than simply a location and scale parameter) and $\gamma$ becomes an index for evaluating the skewness of the SN.

@@

Since the CCF presents a natural asymmetry due the convective blueshift, the SN distribution should in principle better catch this aspect, together with other changes in asymmetry, respect the Normal fit. To initially check this intuition, we compared the CCF residuals after fitting a Normal and a SN distribution for 2 stars. The first star is Alpha Centauri b, whose CCF's have high signal-to-noise ratio (SNR). The second star is Corot-7, whose CCF's have low SNR. Fig.~\ref{fig:Residual.comparison} shows that the SN seems to be a slightly better model to explain the shape of the CCF, in particular as the SNR decreases.
%
\begin{figure*}[htbp]
   \centering
\includegraphics[height = 2.5in]{[1]HD128621Residuals.pdf} 
\includegraphics[height = 2.5in]{[1]LRa01_E2Residuals.pdf} 
   \caption{Comparison between the Normal (black circles) and the SN (red crosses) residuals using CCF's from the star Alpha Centauri b (left) and Corot-7 (right). When looking at the residuals corresponding to the tails of the CCF, the results derived by the two fits are comparable. However, when focusing on the center of the CCF, the SN fit leads to slightly better results for both the stars. Moreover, as the SNR decreases, the SN distribution shows smaller residuals respect the Normal ones.}
    \label{fig:Residual.comparison}
\end{figure*}
%

In the following of the paper, we define RV as the mean of the Normal distribution. Concerning the fit of the CCF using the SN, we present at first $2$ indicators that define the RV of the star: the mean of the SN, defined as SN mean RV and the median of the SN, defined as SN median RV (i.e. realling Eq.\eqref{eq:snmed}, $m$ = SN median RV). We will discuss advantages and limits for both these choices in Sec. \ref{sec:4} and Sec. \ref{sec:5}. For the width of the CCF, we use the FWHM of the Normal, which is $2\sqrt(2\ln2)\sigma$. The width of the SN, SN FWHM, is defined in the same way\footnote{Note that SN FWHM does not correspond to the width of the SN distribution at half maximum like in the Normal case.}. Being a Normal distribution symmetric, there is not such a parameter that evaluates the asymmetry of the distribution, so the BIS SPAN is used. The BIS SPAN will be compared to the asymmetric parameter $\gamma$ of the SN, named SN GAMMA. To test the strength of the correlation between the estimated RV's and the different indicators used to evaluate stellar activity, we calculated the Pearson correlation coefficient, which in its general form is defined as:
%
\begin{equation}
R (x,y)= \frac{cov(x,y)}{\sigma(x),\sigma(y)},
\label{eq:Pearson:corr}
\end{equation}
%
where $x$ and $y$ are two quantitative variables, $cov(x,y)$ indicates the covariance between $x$ and $y$, and $\sigma(x)$ and $\sigma(y)$ represent their standard deviations. 
A $p-$value for the statistical test having null hypothesis $H_{0}: R=0$ is provided, along with a $95\,\%$ confidence interval for $R$.

@@

Because of the spurious variations in RV's caused by active regions, a crucial step of the entire analysis is to correct the initially estimated RV's in order to retrieve a set of RV's containing ideally only the possibly pure doppler shift coming from the exoplanets. In order to reach this goal, we propose to consider a linear combination of the RV's having as covariates the amplitude parameter A (also known as contrast), the BIS SPAN (or $\gamma$), the FWHM (or SN FWHM) and the interaction between the BIS SPAN and the FWHM (or $\gamma$ and SN FWHM and the interaction between $\gamma$ and SN FWHM in the SN case). The parameters in parenthesis will be used when the SN fit is used. We propose the following function to correct the RV's from stellar activity:
%
\begin{equation}
RV_{\text{stellar activity}}= \beta_{0} + \beta_{1} A + \beta_{2} \gamma + \beta_{3} \text{SN FWHM} + \beta_{4} (\gamma  \text{SN FWHM})+\epsilon,
\label{eq:RV:correction}
\end{equation}
%
where $\beta_{0}$ is the intercept and $\epsilon$ is the vector of the errors with mean equal to $0$ and covariance matrix equal to $\sigma^{2}I$ ($I$ defined as the identity matrix).
